\section{Overview}\label{sec:overview}

As the world gets more instrumented and connected, we are witnessing a flood of digital data that is getting generated, in a high velocity, from different hardware (e.g., sensors) or software in the format of streams of data. Examples of this phenomena are crucial for several
applications and domains including financial markets, surveillance systems, manufacturing, smart cities and scalable monitoring infrastructure. In these applications and domains, there is a crucial requirement to collect, process, and analyze big streams of data in order to extract valuable information, discover new insights in real-time and to detect emerging patterns and outliers. Recently, several systems (e.g., \texttt{Apache Apex}\footnote{\url{https://apex.apache.org/}}, \texttt{Apache Flink}\footnote{\url{https://flink.apache.org/}}, \texttt{Apache Storm}\footnote{\url{http://storm.apache.org/}}, \texttt{Heron}\footnote{\url{https://twitter.github.io/heron/}}, \texttt{Spark Streaming}\footnote{\url{https://spark.apache.org/streaming/}}) have been introduced to tackle the real-time processing of big streaming data. However, there are several challenges and open problems that need to be addressed in order improve the state-of-the-art in this domain and push big stream processing systems to make them widely used by large number of users and enterprises~\cite{sakr2016big}.

This report is based on a five-day workshop on "\emph{Big Stream Processing Systems}" that took place in Dagstuhl Schloss  in Germany during the period between 29 October and 3 November 2017. The workshop was attended by 29 researchers from  from 13 countries. Participants of the seminars came from different communities including systems, query languages,
benchmarking, stream mining and semantic stream processing. A major benefit of this workshop was the opportunity for scholars from different communities to
get  exposure to each other and get freely engaged in direct and interactive discussions. The seminar program consisted of different activities including  tutorials that cover the main topics of the workshop and two working groups that have been dedicated to deeply investigating selected topics. The focus of the first group was on applications and system of big stream processing while the second group focused on analyzing the state-of-the-art of stream processing languages. In addition, there were sessions
in which the attendee gave lightening talks on
their own ongoing research work and open research challenges facing them. Highlights about the presentations of the workshop program will be presented in this report.








